Bullet Data Preparation for LIME Diagnostics Paper

### 2020-09-16 README.Rmd
updated data files in preparation for training the random forest model

- Paper
  - Figure out discrepancies with old and new versions of the training data
  - Rerun the lime explanations with the finalized training data
  - Update all files in the submission folder
- Submission
  - Add Heike's ORCID ID
  - Not required but "Authors are requested to suggest at least one recommended reviewer"
  - Edit cover letter and upload
  - Enter the name of the Editor you are submitting to.
  - Edit dataset descriptions in README and add to the readme.txt document for submission
  - Upload the files for submission and submit
- Look into using renev

# LIME Diagnostics Paper

@@

This repository contains the files associated with the paper "Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations" by Katherine Goode and Heike Hofmann. The rest of this readme contains:

- information about the files in the repository,
- instructions for knitting the paper,
- descriptions of the bullet data used in the paper, and
- information on the font sizes used for the figures in the paper.

@@

Contains the R code associated with the paper that is not directly included in `paper.Rnw` (such as some external data cleaning and file organization):

- `data-preparation.Rmd` contains R code for preparing the raw bullet data (bullet training data, bullet testing data, and example matching signatures) to be used in the paper and saves zip versions of the data sets to be uploaded to GitHub
- `figure-compression.Rmd` contains R code for compressing the static figure used in the paper to be uploaded to GitHub
- `submission-steps.Rmd` contains the to do list and R code for preparing the files for submission to the ASA Data Science Journal.

@@

Contains the data used to generate the paper:

- `bullet-test.csv.zip` and `bullet-test.csv.zip` contain the data used for the bullet examples in the paper. Additional information about these files is included below under the heading "Additional Information on Bullet Data".
- `example-signatures.csv.zip` contains signature data from two matching bullet lands that is used to generate a figure in the paper.
- `raw` contains the raw versions of the data files (from Heike and CSAFE) that the above datasets are derived from.

@@

- `figure-08-1.png.zip` is the one figure in the paper that is not generated using R code.
- `figure-readme.png` is a figure used in the readme.

@@

Contains the files associated with an old version of the paper that is much longer the the submitted version

@@

- `.gitignore` contains objects to be ignored by git.
- `paper.Rnw`, `paper.pdf`, and `paper.tex` are the paper. Note that `paper.pdf` and `paper.tex` are generated by `paper.Rnw`.
- `README.Rmd` and `README.md` are the readme files that generated this output.
- `readme.txt` contains descriptions of the data files that are included with the submission
- `references.bib` contains the bibtex information for the references in the paper
- `WileyNJG-AMS.bst` and `WileyNJG-v2.cls` contain the latex formatting specifications for the ASA Data Science Journal.

## Instructions for Knitting Paper

When the paper is knit, various files will be generated (such as `data-bullet-explain.rds` and `figure-10-1.png`). The first time the paper is knit, it will take a while (approximately 45 minutes on Katherine's computer). Once these files have been generated during the first knit, the paper will take much less time to knit (approximately 1 minute on Katherine's computer).

Follow these instructions to knit the paper:

1. Pull the repository from GitHub.
2. Install R packages [goodekat/lime](https://github.com/goodekat/lime) and [goodekat/limeaid](https://github.com/goodekat/limeaid) from GitHub using the remotes R package. (Note that goodekat/lime is a forked versions of [thomasp85/lime](https://github.com/thomasp85/lime) with minor changes to extract objects for the analysis in the paper.)

@@

3. Knit paper (via the paper.Rnw file).

@@

The data used in the paper for the bullet comparison examples is described below.

@@

A dataset with `r dim(bullet_train)[1]` rows and `r dim(bullet_train)[2]` columns that contains comparison features from the Hamby 173 and 252 sets (XXX reference XXX). This dataset is created using the data available at XXX, which contains XXX (add more info about the comparisons). The only changes made to the data from XXX were to remove any lands with tank rash, select only variables of use for the paper, add a case variable, rename `same_source` as `samesource`, and attach the random forest scores associated with the comparisons. The variables in the data are as follows:

- `r names(bullet_train)[1]` = ID number associated with the bullet-land signature comparison
- `r names(bullet_train)[2]` = ID associated with the first bullet-land in the comparison. The format is 'study - barrel number - bullet number - land number'
- `r names(bullet_train)[3]` = ID associated with the second bullet-land in the comparison. Again, the format is 'study - barrel number - bullet number - land number'
- `r names(bullet_train)[4]`, `r names(bullet_train)[5]`, `r names(bullet_train)[6]`, `r names(bullet_train)[7]`, `r names(bullet_train)[8]`, `r names(bullet_train)[9]`, `r names(bullet_train)[10]`, `r names(bullet_train)[11]`, and `r names(bullet_train)[12]` = comparison features of the two bullet-land signatures. See XXX for definitions of these features
- `r names(bullet_train)[13]` = indicator variable specifying whether the two bullet-lands are a match
- `r names(bullet_train)[14]` = the random forest score associated with the comparison in support of a match

@@

- `r names(bullet_test)[1]` = ID number associated with the bullet-land signature comparison
- `r names(bullet_test)[2]` = ID associated with the first bullet-land in the comparison with a format of 'study - set number - bullet number - land number'
- `r names(bullet_test)[3]` = ID associated with the second bullet-land in the comparison with a format of 'study - set number - bullet number - land number'
- `r names(bullet_test)[4]`, `r names(bullet_test)[5]`, `r names(bullet_test)[6]`, `r names(bullet_test)[7]`, `r names(bullet_test)[8]`, `r names(bullet_test)[9]`, `r names(bullet_test)[10]`, `r names(bullet_test)[11]`, and `r names(bullet_test)[12]` = comparison features of the two bullet-land signatures. See XXX for definitions of these features.
- `r names(bullet_test)[13]` = indicator variable specifying whether the two bullet-lands are a match
- `r names(bullet_test)[14]` = the random forest score associated with the comparison in support of a match

@@

The font sizes in the figures are calculated based on how the figures are scaled due to the specification of both fig.width and out.width in the Rnw file to ensure that they are consistent across figures. The text sizes are either set to 7 pt or 5.5 pt (Helvetica) as indicated by the table below. Note that the text in the paper is 9 pt (Times) and a similar approach was used to ensure that line sizes are larger than 0.5 pt.

>>>>

- Manuscript
  - Adjust order of match and non-match in feature heat map
  - Train a random forest model on prepared training data (make sure to set a seed!) 
  - Make predictions on test data
  - Rerun all code in manuscript (check for changes in results)
  - Update text in manuscript to describe the random forest model process
  - Update all files in the submission folder
- Submission
  - Edit dataset descriptions in README and add to the readme.txt document for submission
  - Upload the files for submission and submit

# LIME Diagnostics Manuscript

@@

This repository contains the files associated with the manuscript "Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations" by Katherine Goode and Heike Hofmann. The rest of this readme contains:

- information about the files in the repository,
- instructions for knitting the manuscript,
- descriptions of the bullet data used in the manuscript, and
- information on the font sizes used for the figures in the manuscript.

@@

Contains the R code associated with the manuscript that is not directly included in `paper.Rnw` (such as some external data cleaning and file organization):

- `data-preparation.Rmd` contains R code for preparing the raw bullet data (bullet training data, bullet testing data, and example matching signatures) to be used in the manuscript and saves zip versions of the data sets to be uploaded to GitHub
- `figure-compression.Rmd` contains R code for compressing the static figure used in the manuscript to be uploaded to GitHub
- `submission-steps.Rmd` contains the to do list and R code for preparing the files for submission to the ASA Data Science Journal.

@@

Contains the data used to generate the manuscript:

- `bullet-test.csv.zip` and `bullet-test.csv.zip` contain the data used for the bullet examples in the manuscript. Additional information about these files is included below under the heading "Additional Information on Bullet Data".
- `example-signatures.csv.zip` contains signature data from two matching bullet lands that is used to generate a figure in the manuscript.
- `raw` contains the raw versions of the data files (from Heike and CSAFE) that the above datasets are derived from.

@@

- `figure-08-1.png.zip` is the one figure in the manuscript that is not generated using R code.
- `figure-readme.png` is a figure used in the readme.

@@

Contains the files associated with an old version of the manuscript that is much longer the the submitted version

@@

- `.gitignore` contains objects to be ignored by git.
- `paper.Rnw`, `paper.pdf`, and `paper.tex` are the manuscript. Note that `paper.pdf` and `paper.tex` are generated by `paper.Rnw`.
- `README.Rmd` and `README.md` are the readme files that generated this output.
- `readme.txt` contains descriptions of the data files that are included with the submission
- `references.bib` contains the bibtex information for the references in the manuscript
- `WileyNJG-AMS.bst` and `WileyNJG-v2.cls` contain the latex formatting specifications for the ASA Data Science Journal.

## Instructions for Knitting Manuscript

When the manuscript is knit, various files will be generated (such as `data-bullet-explain.rds` and `figure-10-1.png`). The first time the manuscript is knit, it will take a while (approximately 45 minutes on Katherine's computer). Once these files have been generated during the first knit, the manuscript will take much less time to knit (approximately 1 minute on Katherine's computer).

Follow these instructions to knit the manuscript:

1. Pull the repository from GitHub.
2. Install R packages [goodekat/lime](https://github.com/goodekat/lime) and [goodekat/limeaid](https://github.com/goodekat/limeaid) from GitHub using the remotes R package. (Note that goodekat/lime is a forked versions of [thomasp85/lime](https://github.com/thomasp85/lime) with minor changes to extract objects for the analysis in the manuscript.)

@@

3. Knit manuscript (via the `paper.Rnw` file).

@@

The data used in the manuscript for the bullet comparison examples is described below.

@@

A dataset with `r dim(bullet_train)[1]` rows and `r dim(bullet_train)[2]` columns that contains comparison features from the Hamby 173 and 252 sets (XXX reference XXX). It contains comparisons from `r length(unique(bullet_train$land_id1))` bullet-land signatures with some lands missing due to tank rash or other reasons. This dataset is created using the data available at XXX, which contains XXX (add more info about the comparisons). The only changes made to the data from XXX were to remove any lands with tank rash, select only variables of use for the manuscript, add a case variable, rename `same_source` as `samesource`, and attach the random forest scores associated with the comparisons. The variables in the data are as follows:

- `r names(bullet_train)[1]` = ID number associated with the bullet-land signature comparison
- `r names(bullet_train)[2]` = ID associated with the first bullet-land in the comparison. The format is 'study - barrel number - bullet number - land number'
- `r names(bullet_train)[3]` = ID associated with the second bullet-land in the comparison. Again, the format is 'study - barrel number - bullet number - land number'
- `r names(bullet_train)[4]`, `r names(bullet_train)[5]`, `r names(bullet_train)[6]`, `r names(bullet_train)[7]`, `r names(bullet_train)[8]`, `r names(bullet_train)[9]`, `r names(bullet_train)[10]`, `r names(bullet_train)[11]`, and `r names(bullet_train)[12]` = comparison features of the two bullet-land signatures. See XXX for definitions of these features
- `r names(bullet_train)[13]` = indicator variable specifying whether the two bullet-lands are a match

@@

- `r names(bullet_test)[1]` = ID number associated with the bullet-land signature comparison
- `r names(bullet_test)[2]` = ID associated with the first bullet-land in the comparison with a format of 'study - set number - bullet number - land number'
- `r names(bullet_test)[3]` = ID associated with the second bullet-land in the comparison with a format of 'study - set number - bullet number - land number'
- `r names(bullet_test)[4]`, `r names(bullet_test)[5]`, `r names(bullet_test)[6]`, `r names(bullet_test)[7]`, `r names(bullet_test)[8]`, `r names(bullet_test)[9]`, `r names(bullet_test)[10]`, `r names(bullet_test)[11]`, and `r names(bullet_test)[12]` = comparison features of the two bullet-land signatures. See XXX for definitions of these features.
- `r names(bullet_test)[13]` = indicator variable specifying whether the two bullet-lands are a match

@@

The font sizes in the figures are calculated based on how the figures are scaled due to the specification of both fig.width and out.width in the Rnw file to ensure that they are consistent across figures. The text sizes are either set to 7 pt or 5.5 pt (Helvetica) as indicated by the table below. Note that the text in the manuscript is 9 pt (Times) and a similar approach was used to ensure that line sizes are larger than 0.5 pt.

### 2020-09-16 code/data-preparation.Rmd
added a missed r package
updated data files in preparation for training the random forest model

```{r}
library(cowplot)
library(dplyr)
library(forcats)
library(ggplot2)
library(randomForest)
library(stringr)
library(tidyr)
```

@@

Create a new version of the data with the land IDs separated into separate variables for an analysis of the number of observations in this data and reorder the barrels:

```{r}
bullet_train_raw_expanded <-
  bullet_train_raw %>%
  separate(land_id1,
           sep = "-",
           into = c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(land_id2,
           sep = "-",
           into = c("study2", "barrel2", "bullet2", "land2")) %>%
  mutate(barrel1 = fct_relevel(barrel1, paste0("Br", 1:10), "BrUnk"),
         barrel2 = fct_relevel(barrel2, paste0("Br", 1:10), "BrUnk"))
```

The number of comparisons within the two studies are included below:

```{r}
table(
  bullet_train_raw_expanded$study1,
  bullet_train_raw_expanded$study2
)
```

@@

**XXX Need to determine why the number of dimensions do not match. XXX**

@@

Add a variable to the training data with the random forest probability in support of a match using the random forest model `rtrees` from `bulletxtrctr`:

```{r}
bullet_features <- rownames(bulletxtrctr::rtrees$importance)
bullet_train$rfscore <- 
  bulletxtrctr::rtrees %>%
  predict(bullet_train %>% select(all_of(bullet_features)), type = "prob") %>%
  data.frame() %>%
  pull(TRUE.)
```

@@

```{r}
hamby224_set1_cleaned <-
  hamby224_set1 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set1"),
    bullet1 = recode(factor(bullet1), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    bullet2 = recode(factor(bullet2), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    land1 = paste0("L", land1),
    land2 = paste0("L", land2)
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rf_features), rfscore, samesource)

hamby224_set11_cleaned <- 
  hamby224_set11 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set11"),
    bullet1 = recode(factor(bullet1), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    bullet2 = recode(factor(bullet2), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    land1 = str_remove(land1, "and "),
    land2 = str_remove(land2, "and ")
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rf_features), rfscore, samesource)
```

@@

# Select and order variables to keep for analysis
hamby224_test <- hamby224_test_extra_vars %>%
  select(case, land_id1, land_id2, all_of(rf_features), samesource, rfscore)
```

>>>>

```{r}
library(cowplot)
library(dplyr)
library(forcats)
library(ggplot2)
library(randomForest)
library(stringr)
library(tidyr)
library(zip)
```

@@

Determine the number of total study-barrel-bullet-lands in the data:

```{r}
length(unique(bullet_train_raw %>% pull(land_id1)))
```

Determine the number of total study-barrel-bullet-lands in the data when observations with tank rash are excluded:

```{r}
length(unique(bullet_train_raw %>% filter(flag == FALSE) %>% pull(land_id1)))
```

There is only one less land id when the cases with tank rash are excluded. The code below shows that this case is `Hamby252-Br9-B2-L4` since there are no rows in the filtered data without `Hamby252-Br9-B2-L4` as one of the land ids:

```{r}
length(unique(
  bullet_train_raw %>% filter(flag == TRUE) %>% filter(
    land_id1 != "Hamby252-Br9-B2-L4",
    land_id2 != "Hamby252-Br9-B2-L4"
  ) %>% pull(land_id1)
))
```

@@

The discrepancy in the number of observations has led us to decide to train a new random forest model using the data provided by Heike where `flag == FALSE` (observations known to not have tank rash). The code for the random forest can be found in the document `paper.Rnw`.

@@

@@

```{r}
hamby224_set1_cleaned <-
  hamby224_set1 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features,-rfscore) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set1"),
    bullet1 = recode(factor(bullet1), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    bullet2 = recode(factor(bullet2), "1" = "B1", "2" = "B2", "Q" = "BUnk"),
    land1 = paste0("L", land1),
    land2 = paste0("L", land2)
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rf_features), samesource)

hamby224_set11_cleaned <- 
  hamby224_set11 %>%
  # Remove unneeded variables
  select(-bullet_score,-land1,-land2,-aligned,-striae,-features,-rfscore) %>%
  # Rename bullet and land variables
  rename(
    bullet1 = bulletA,
    bullet2 = bulletB,
    land1 = landA,
    land2 = landB
  ) %>%
  # Create study and set variables and change levels of bullet and land
  # variables to match those in the training data
  mutate(
    study = factor("Hamby224"),
    set = factor("Set11"),
    bullet1 = recode(factor(bullet1), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    bullet2 = recode(factor(bullet2), "Bullet 1" = "B1", "Bullet 2" = "B2", "Bullet I" = "BUnk"),
    land1 = str_remove(land1, "and "),
    land2 = str_remove(land2, "and ")
  ) %>%
  # Order the variables as desired
  select(study, set, bullet1, land1, bullet2, land2, all_of(rf_features), samesource)
```

@@

# Select and order variables to keep for analysis
hamby224_test <- hamby224_test_extra_vars %>%
  select(case, land_id1, land_id2, all_of(rf_features), samesource)
```
